---
title: "STA_445_Assignment_6"
author: "Gabriel Cage-Sepeda"
date: "2024-3-26"
output: pdf_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
```

## Problem 1
Convert the following to date or date/time objects.

a. September 13, 2010.

```{r}
mdy('September 13, 2010')
```

b. Sept 13, 2010.

```{r}
mdy('Sept 13, 2010')
```

c. Sep 13, 2010.

```{r}
mdy('Sep 13, 2010')
```

d. S 13, 2010. Comment on the month abbreviation needs.

#The month needs the corrct abbreviation, for September only Sep works as an abbreviation.

```{r}
mdy('S 13, 2010')
```

e. 07-Dec-1941.

```{r}
dmy('07-Dec-1941')
```

f. 1-5-1998. Comment on why you might be wrong.

#1 and 5 could both be the month or day, so this is either May first or January fifth. I went with May first.

```{r}
dmy('1-5-1998')
```

g. 21-5-1998. Comment on why you know you are correct.

#21 can't be the month and 1998 has to be the year, so this is May 21.

```{r}
dmy('21-5-1998')
```

h. 2020-May-5 10:30 am

```{r}
ymd_hm('2020-May-5 10:30 am')
```

i. 2020-May-5 10:30 am PDT (ex Seattle)

```{r}
ymd_hm('2020-May-5 10:30 am', tz = 'US/Pacific')
```

j. 2020-May-5 10:30 am AST (ex Puerto Rico)

```{r}
ymd_hm('2020-May-5 10:30 am', tz = 'Atlantic/Bermuda')
```

### Problem 2

Using just your date of birth (ex Sep 7, 1998) and today's date calculate the following:

```{r}
bday <- mdy_hm('12-28-2003 12:30pm', tz = 'US/Arizona')
current <- mdy_hm('03-26-2024 3:00pm', tz = 'US/Arizona')
bday
current
```

a. Calculate the date of your 64th birthday.

```{r}
bday + years(64)
```

b. Calculate your current age (in years).

```{r}
year(as.period(interval(bday, current)))
```

c. Using your result in part (b), calculate the date of your next birthday.

```{r}
Nextbday <- bday + years(20) + years(1)
Nextbday
```

d. The number of _days_ until your next birthday.

```{r}
as.period(interval(current, Nextbday), unit = 'days')
```

e. The number of _months_ and _days_ until your next birthday.

```{r}
as.period(interval(current, Nextbday))
```

## Problem 3

Suppose you have arranged for a phone call to be at 3 pm on May 8, 2015 at Arizona time. However, the recipient will be in Auckland, NZ. What time will it be there? 

```{r}
with_tz(mdy_hm('05-08-2015 3:00 pm', tz = 'US/Arizona'), tzone = 'Pacific/Auckland')
```

## Problem 4

It turns out there is some interesting periodicity regarding the number of births on particular days of the year.

a. Using the `mosaicData` package, load the data set `Births78` which records the number of children born on each day in the United States in 1978. Because this problem is intended to show how to calculate the information using the `date`, remove all the columns *except* `date` and `births`. 

```{r warning=FALSE}
library(mosaicData)
data(Births78)
Births78 <- Births78 %>%
  select('date', 'births')
Births78
```

b. Graph the number of `births` vs the `date` with date on the x-axis. What stands out to you? Why do you think we have this trend?

#There are more births around August and September than the rest of the year, maybe because the holiday season is 9 months before. There is also a space between clusters of dots in each month, proabably for day of the week.

```{r}
Births78 %>%
  ggplot(aes(x=date, y=births)) +
  geom_point()
```

c. To test your assumption, we need to figure out the what day of the week each observation is. Use `dplyr::mutate` to add a new column named `dow` that is the day of the week (Monday, Tuesday, etc). This calculation will involve some function in the `lubridate` package and the `date` column.

```{r}
Births78 <- Births78 %>%
  dplyr::mutate('dow' = wday(date, label=TRUE))
Births78
```

d. Plot the data with the point color being determined by the day of the week variable.

#The space between clusters in each month is because there are less births on the weekend (Saturday/Sunday) than during thw week.

```{r}
Births78 %>%
  ggplot(aes(x=date, y=births)) +
  geom_point(aes(color=dow))
```
