---
title: "Assign. 1 STA 445"
author: "Gabriel Cage-Sepeda"
date: "`r Sys.Date()`"
output: pdf_document
---

## Directions:
This assignment covers chapter 5.  Please show all work in this document and knit your final draft into a pdf.  This is assignment is about statistical models, which will be helpful if you plan on taking STA 570, STA 371, or STA 571. 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## \color{black} Problem 1: Two Sample t-test

\color{blue} 

a. Load the `iris` dataset.

```{r}
data('iris')
```

b. Create a subset of the data that just contains rows  for the two species setosa and versicolor using filter.  Use slice_sample to print out 20 random rows of the dataset.

```{r}
iris1 <- iris %>%
  filter(Species=="setosa" | Species=="versicolor")
slice_sample(iris1)
```

c. Create a box plot of the petal lengths for these two species using ggplot. Does it look like the mean petal length varies by species?

```{r}
iris1 %>%
  ggplot(aes(x=Species, y=Petal.Length)) +
  geom_boxplot()
```

d. Do a two sample t-test using t.test to determine formally if the petal lengths differ. Note:  The book uses the tidy function in the broom package to make the output "nice".  I hate it!  Please don't use tidy.

```{r}
t.test(data=iris1, Petal.Length ~ Species)
```

d. What is the p-value for the test? What do you conclude?

P-value=2.2x10^-16

Since the p-value is very small, reject the null hypothesis.

There is significant evidence of a difference in mean petal length between setosa and versicolor flowers.
 
e. Give a 95% confidence interval for the difference in the mean petal lengths.

(-2.939618, -2.656382)

f. Give a 99% confidence interval for the difference in mean petal lengths.  (Hint: type ?t.test.  See that you can change the confidence level using the option conf.level)

```{r}
t.test(data=iris1, Petal.Length ~ Species, conf.level=.99)
```
99% confidence level: (-2.986265, -2.609735)

g. What is the mean petal length for setosa?

1.462

h. What is the mean petal length for versicolor?

4.260

## Problem 2: ANOVA 

Use the iris data with all three species.

a. Create a box plot of the petal lengths for all three species using ggplot.Does it look like there are differences in the mean petal lengths?

```{r}
iris %>%
  ggplot(aes(x=Species, y=Petal.Length)) +
  geom_boxplot()
```

b. Create a linear model where sepal length is modeled by species. Give it
an appropriate name.

```{r}
iris.model <- lm(data=iris, Sepal.Length ~ Species-1)
summary(iris.model)
```

c. Type anova(your model name) in a code chunk.

```{r}
anova(iris.model)
```

d. What is the p-value for the test?  What do you conclude.

p-value= 2.2x10^-16.

Since p is very small, reject the null.

There is significant evidence that sepal length differs by species of flower.

e. Type summary(your model name) in a code chunk.

```{r}
summary(iris.model)
```

f. What is the mean sepal length for the species setosa?

5.0060

g. What is the mean sepal length for the species versicolor?

5.9360

## Problem 3: Regression 

Can we describe the relationship between petal length and petal width?

a. Create a scatterplot with petal length on the y-axis and petal width on the x-axis using ggplot.

```{r}
ggplot(data=iris, aes(x=Petal.Width, y=Petal.Length)) +
  geom_point()
```

b. Create a linear model to model petal length with petal width (length is the response variable and width is the explanatory variable) using lm.

```{r}
petal.model <- lm(data=iris, Petal.Length ~ Petal.Width)
summary(petal.model)
```

c. What is the estimate of the slope parameter?

2.22994

d. What is the estimate of the intercept parameter?

1.08356

e. Use summary() to get additional information.

```{r}
summary(petal.model)
```

## Problem 4: Modeling Trees

Using the `trees` data frame that comes pre-installed in R, follow the steps below to fit the regression model that uses the tree `Height` to explain the `Volume` of wood harvested from the tree.

a. Create a scatterplot of the data using ggplot.

```{r}
ggplot(data=trees, aes(x=Height, y=Volume)) +
  geom_point()
```

b. Fit a `lm` model using the command `model <- lm(Volume ~ Height, data=trees)`.

```{r}
model <- lm(Volume ~ Height, data=trees)
```

c. Print out the table of coefficients with estimate names, estimated value, standard error, and upper and lower 95% confidence intervals.

```{r}
summary(model)
```

d.Add the model fitted values to the `trees` data frame along with the regression model confidence intervals.  Note: the book does this in a super convoluted way.  Don't follow the model in the book.  Instead try cbind.

```{r}
predict.model <- cbind(trees, predict(model, interval = "confidence"))

predict.model
```

e. Graph the data and fitted regression line and uncertainty ribbon.

```{r}
ggplot(data=predict.model, aes(x=Height, y=Volume)) +
  geom_point() +
  geom_line(aes(y=fit)) +
  geom_ribbon(aes(ymin=lwr, ymax=upr), alpha=0.5, fill="skyblue")
```

f. Add the R-squared value as an annotation to the graph using annotate.

```{r}
ggplot(data=predict.model, aes(x=Height, y=Volume)) +
  geom_point() +
  geom_line(aes(y=fit)) +
  geom_ribbon(aes(ymin=lwr, ymax=upr), alpha=0.5, fill="skyblue") +
  annotate("text", x=75, y=75, label="R-squared = 0.3579")
```

